Blaming the Typeless: \\ Scalable, Human-Centric Python Localization

### 2018-08-24 paper/main.tex
graphs
paper

\paragraph{Training methodology}
At this point we have feature vectors describing every term of every program in
our data set. The vectors from a random 30\% of the program pairs are set aside for testing and
the other 70\% are used for training.
Within the training data, we use 3-fold
cross-validation to select the best parameters for our models, before training
using those parameters on the full training set and then scoring the models
using the testing set.
Each trained model takes in any vector representing a term
in a buggy program, and returns a confidence score representing how likely it is that
that term was one of the terms changed between the fixed and buggy versions.
Thus we can treat the model as providing a \emph{ranking} over all terms by
confidence, where the top result is the one that the model deems most likely
to be changed.
For a given $k$, we score the model based on Top-$k$ accuracy: for what
proportion of the programs is a correct answer (i.e. a term that is actually
changed) present in the top $k$ results.
We retrain the models for each value of k, since a model optimized to produce
the single best result might be different from one optimized to get a good
result into the top three.

For decision tree and random forest, we use the implementation from
\textsc{scikit-learn}~\cite{scikit-learn}.
Due to the large amount of data points, we did not use \textsc{scikit-learn}'s
implementation for logistic regression and multilayer perceptron.
We implementation both model using
\textsc{keras}~\cite{chollet2015keras} with mini-batch gradient descent.

Because of the imbalance nature of the data set, where not-blamed lines
(negative examples) appears more frequently than blamed lines (positive
examples).
The weight for each example during training is re-weighted to the reciprocal
of the frequency of corresponding class.

\paragraph{Hyper-parameters}
All hyper-parameters are tuned using three-fold cross-validation on the
intended score (Top-$k$ accuracy).
For RF, the maximum depth of each tree is searched from $25$ to $35$ and the
number of tree are fixed to $500$.
For decision tree maximum depth of the tree is searched from $25$ to $35$ with interval of $5$,
the minimum impurity decrease for each split is searched from $10^{-7}$ to
$10^{-3}$ with multiple of $10^{-2}$ and the minimum samples in each leaf is set to $200$.
For logistic regression and MLP, binary cross-entropy is set as the loss
function and Adam \cite{kingma2014adam} optimizer is used.
Learning rate is set to $0.1$ and the number of epochs are $10$.
L2 regularization is applied to the weights of both models,
the weight for L2 term is searched from $10^0$ to $10^{-5}$ with a multiple of $10^{-1}$.
Other parameters that are not mentioned are kept with the default setting of the
original packages.

@@

\item[RQ3]{Which features are the most important for various categories of
Python defects?}

@@

\begin{figure}

\caption{\fixme{This figure has been removed; this is a placeholder so the
numbering doesn't change while I handle some TODOs}}
\label{fig-full-dataset-acc}
\end{figure}

@@

We first use the full dataset to train and test decsion forests optimized for
Top-1, Top-2, and Top-3 accuracy. As shown in \figref{full-dataset-acc},
these produce a correct answer 60, 72, and
79\% of the time (to two significant figures). As a baseline, we determine that
the expression blamed by the standard Python interpreter (i.e. the one that
causes it to crash) is only changed by the user 30\% of the time. Thus our
most directly comparable model, the one for Top-1, significantly outperforms
this baseline, and as we then exercise our power to give a ranked output rather
than a single answer, our margin gets even better.

@@

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={No Context, No Types},
    xtick=data,
    ymin=0,
    ymax=1.1,
    ylabel=Normalized Model Accuracy,
    enlarge x limits=0.5,
    legend style={at={(0.5,-0.15)},anchor=north}
]
    \addplot table[x=scoreName, y=score, col sep=comma]{removing-features-2-normalized.csv};
    \draw [red] ({rel axis cs:0,0}|-{axis cs:No Types,1}) -- ({rel axis cs:1,0}|-{axis cs:No Types,1}) node [pos=0.33, above] {};
\end{axis}
\end{tikzpicture}
\caption{Accuracy when features are removed, based on Top-1 decision trees on a
random subset of 20000 program pairs, as compared to a Top-1 decision tree trained
on all features.
}
\label{fig-removing-features}
\end{figure}

@@

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={Index, Value, Key, Type, Attribute},
    xtick=data,
    ymin=0,
    ylabel=Normalized Model Accuracy,
    legend style={at={(0.5,-0.15)},anchor=north}
]
    \addplot table[x=scoreName, y=score, col sep=comma]{defect-categories-2-normalized.csv};
    \draw [red] ({rel axis cs:0,0}|-{axis cs:Index,1}) -- ({rel axis cs:1,0}|-{axis cs:Index,1}) node [pos=0.33, above] {};
\end{axis}
% \begin{axis}
%     \addplot{1};
% \end{axis}
\end{tikzpicture}
\caption{Accuracy when a Top-1 decision tree is trained and tested on only the data exhibiting
the error on the x-axis, as compared to a Top-1 decision tree on all the data.
\fixme{Spin this for generality. Indicate that our model gives high-quality
fault localization regardless of defect type. To put it another way, we
can view this as a sensitivity analysis: our algorithm is not sensitive to
defect type (that's good!).}
}
\label{fig:defect-categories}
\end{figure}

prereq: define defect category. This could be either the raw python
exception name or it could be clusters that we have manually created.
Ranjit notes: result could be ``we need everything for everything'', at
which point this is useless and should be skipped. Per Ben's exam: may find
that for different categories of errors we should use different ML
classifiers.

PLOT Type: Bar Graph (three bars per point on the X axis)

PLOT Y Axis: Leave-One-In Accuracy (normalized per X axis point)
Ranjit notes: on smaller subsets we could rebuild the classifier each time.
Wes notes that if we do, it takes us longer. If we re-use the monolithic
model we can do this very quickly.

PLOT X AXis: Defect Categories

PLOT Bars: Static, Dynamic, Context

Goal: Demonstrate that each category of feature is essential for a certain
piece of the problem. We need ABC feature to handle DEF class of defects,
but ABC feature does not work well on GHI class of defects, for those we
need JKL features. Implicit: we were smart for including all of these
categories.

@@

\subsection{RQ 5 --- Wishlist: Partitioning Kinds of Errors}

Can we partition the kinds of errors that people make? What are common
misconceptions? Do these correspond to paths of decision trees.

If we don't get to the source-sink thing, it could go here.

I don't think we have time, in this paper, to actually implement a
two-level machine learning.

>>>>

\emph{Training methodology}
At this point we have feature vectors describing every term of every program in
our data set. The vectors from a random 30\% of the program pairs are set aside for testing and
the other 70\% are used for training.
Within the training data, we use 3-fold
cross-validation to select the best parameters for our models, before training
using those parameters on the full training set and then scoring the models
using the testing set.
Each trained model takes in any vector representing a term
in a buggy program, and returns a confidence score representing how likely it is that
that term was one of the terms changed between the fixed and buggy versions.
Thus we can treat the model as providing a \emph{ranking} over all terms by
confidence, where the top result is the one that the model deems most likely
to be changed.
For a given $k$, we score the model based on Top-$k$ accuracy: for what
proportion of the programs is a correct answer (i.e. a term that is actually
changed) present in the top $k$ results.
We retrain the models for each value of k, since a model optimized to produce
the single best result might be different from one optimized to get a good
result into the top three.

For decision tree and random forest, we use the implementation from
\textsc{scikit-learn}~\cite{scikit-learn}.
% Due to the large amount of data points, we did not use \textsc{scikit-learn}'s
% implementation for logistic regression and multilayer perceptron.
% We implementation both model using
% \textsc{keras}~\cite{chollet2015keras} with mini-batch gradient descent.

Unsurprisingly this is an imbalanced dataset in that negative examples, i.e.
terms that are not changed by the programmer, are much more common than
positive examples, so during training we re-weight each example to the reciprocal
of the frequency of corresponding class.

\emph{Hyper-parameters}
All hyper-parameters are tuned using three-fold cross-validation on the
intended score (Top-$k$ accuracy).
For RF, the maximum depth of each tree is searched from $25$ to $35$ and the
number of tree are fixed to $500$.
For decision tree maximum depth of the tree is searched from $25$ to $35$ with interval of $5$,
the minimum impurity decrease for each split is searched from $10^{-7}$ to
$10^{-3}$ with multiple of $10^{-2}$ and the minimum samples in each leaf is set to $200$.
For logistic regression and MLP, binary cross-entropy is set as the loss
function and Adam \cite{kingma2014adam} optimizer is used.
Learning rate is set to $0.1$ and the number of epochs are $10$.
L2 regularization is applied to the weights of both models,
the weight for L2 term is searched from $10^0$ to $10^{-5}$ with a multiple of $10^{-1}$.
Other parameters that are not mentioned are kept with the default setting of the
original packages.

@@

% \item[RQ3]{Which features are the most important for various categories of
% Python defects?}

\item[RQ3]{How well does our algorithm handle different kinds of Python errors?}

@@

@@

We first use the full dataset to train and test decsion forests optimized for
Top-1, Top-2, and Top-3 accuracy. As shown in \figref{full-dataset-acc-random-forest},
these produce a correct answer 60, 72, and
79\% of the time (to two significant figures). As a baseline, we determine that
the expression blamed by the standard Python interpreter (i.e. the one that
causes it to crash) is only changed by the user 45\% of the time. Thus our
most directly comparable model, the one for Top-1, significantly outperforms
this baseline, and as we then exercise our power to give a ranked output rather
than a single answer, our margin gets even better.

@@

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={No Context, No Types},
    xtick=data,
    ymin=0,
    ymax=1.1,
    ylabel=Normalized Model Accuracy,
    enlarge x limits=0.5,
    legend style={at={(0.5,-0.15)},anchor=north}
]
    \addplot table[x=scoreName, y=score, col sep=comma]{removing-features-2-normalized.csv};
    \draw [red] ({rel axis cs:0,0}|-{axis cs:No Types,1}) -- ({rel axis cs:1,0}|-{axis cs:No Types,1}) node [pos=0.33, above] {};
\end{axis}
\end{tikzpicture}
\caption{Accuracy when features are removed, based on Top-1 decision trees on a
random subset of 20000 program pairs, as compared to a Top-1 decision tree trained
on all features.
\fixme{will add No Syntax bar}
}
\label{fig-removing-features}
\end{figure}

@@

\begin{figure}
\begin{tikzpicture}
\begin{axis}[
    ybar,
    symbolic x coords={Index, Value, Key, Type, Attribute},
    xtick=data,
    ymin=0,
    ylabel=Normalized Model Accuracy,
    legend style={at={(0.5,-0.15)},anchor=north}
]
    \addplot table[x=scoreName, y=score, col sep=comma]{defect-categories-2-normalized.csv};
    \draw [red] ({rel axis cs:0,0}|-{axis cs:Index,1}) -- ({rel axis cs:1,0}|-{axis cs:Index,1}) node [pos=0.33, above] {};
\end{axis}
\end{tikzpicture}
\caption{Accuracy when a Top-1 decision tree is trained and tested on only the data exhibiting
the error on the x-axis, as compared to a Top-1 decision tree on all the data.
}
\label{fig-defect-categories}
\end{figure}

We next try training and testing Top-1 decision trees on only the subsets of
the data
corresponding to each of the five most common exceptions: \lt{TypeError},
\lt{IndexError}, \lt{AttributeError}, \lt{ValueError}, and \lt{KeyError}.
As shown in \figref{defect-categories}, these models have accuracy between
86\% and 115\% of a comparable model trained on the data as a whole. This shows
that our algorithm is robust, able to give high accuracy fault localizations on
a wide variety of defect types.

% prereq: define defect category. This could be either the raw python
% exception name or it could be clusters that we have manually created.
% Ranjit notes: result could be ``we need everything for everything'', at
% which point this is useless and should be skipped. Per Ben's exam: may find
% that for different categories of errors we should use different ML
% classifiers.

% PLOT Type: Bar Graph (three bars per point on the X axis)

% PLOT Y Axis: Leave-One-In Accuracy (normalized per X axis point)
% Ranjit notes: on smaller subsets we could rebuild the classifier each time.
% Wes notes that if we do, it takes us longer. If we re-use the monolithic
% model we can do this very quickly.

% PLOT X AXis: Defect Categories

% PLOT Bars: Static, Dynamic, Context

% Goal: Demonstrate that each category of feature is essential for a certain
% piece of the problem. We need ABC feature to handle DEF class of defects,
% but ABC feature does not work well on GHI class of defects, for those we
% need JKL features. Implicit: we were smart for including all of these
% categories.

@@

% \subsection{RQ 5 --- Wishlist: Partitioning Kinds of Errors}

% Can we partition the kinds of errors that people make? What are common
% misconceptions? Do these correspond to paths of decision trees.

% If we don't get to the source-sink thing, it could go here.

% I don't think we have time, in this paper, to actually implement a
% two-level machine learning.
